\section{Problem 1 – Automata Theory}

\paragraph{Question 1}
$\Sigma = \{a,b\}$. 
Give NFA $\mathcal{A}_1$ with no more than $4$ states recognizing $L_1 = \{w\in \Sigma \:|\: \exists l\in \Sigma \: |w|_l > 1\}$.
\begin{table}[!h]
\centering
\begin{tabular}{ll||l|l}
  &    & a     & b     \\
  \hline
s & $q_0$ & $q_0$,$q_1$ & $q_0$,$q_2$ \\
  & $q_1$ & $q_3$    & $q_1$    \\
  & $q_2$ & $q_2$    & $q_3$    \\
* & $q_3$ & $q_3$    & $q_3$   
\end{tabular}
\end{table}


\paragraph{Question 2}
Prove: if $\Sigma$ is finite alphabet, then any finite language $L = \{w_1,\cdots,w_n\} \subseteq \Sigma^*$ is regular, $n\in \mathbb{N}$.

That is we should construct a finite automaton accepting $L$.
We can construct a $\epsilon$-NFA containing $n$ "branches", each recognizing $w_i$.
Now, for every $\epsilon$-NFA, there must exist equivalent DFA recognizing the same language.

We can also construct the DFA explicitly.
Start with automaton (NFA) recognizing $w_1$: there are $|w_1|+1$ states, the last one is accepting and transitions are labeled with next letters of $w_1$.
For $w_i$ ($i=2,\cdots,n$) try to traverse the automaton as far as you can, i.e. until transition for symbol $w_{ij}$ exist.
If we can go no further, that is we read the longest common prefix of $w_i$ and some $w_k$, $k<i$, then we make a new branch from a current state.
This branch consist of states and transitions labeled $w_{i,j+1}\cdots w_{i, |w_i|}$.

Now we need to assure that we constructed a DFA.
For every state missing some transitions on some letters, add those transition leading to a "dead state", i.e. nonaccepting state with a self-loop labeled $\Sigma$.

\paragraph{Question 3}
Give DFA recognizing complement of $L_1$ from (Q1), i.e $L_2 = \Sigma^* \setminus L_1$.

Obviously $L_2 = \{w\in \Sigma \:|\: \forall l\in \Sigma \: |w|_l \leq 1\} = \{\epsilon, a, b, ab, ba\}$.
\begin{table}[!h]
\centering
\begin{tabular}{ll||l|l}
    &    & a  & b  \\
    \hline
s,* & $q_1$ & $q_2$ & $q_3$ \\
*   & $q_2$ & $q_5$ & $q_4$ \\
*   & $q_3$ & $q_4$ & $q_5$ \\
*   & $q_4$ & $q_5$ & $q_5$ \\
    & $q_5$ & $q_5$ & $q_5$
\end{tabular}
\end{table}

\paragraph{Question 4}
Given NFA $\mathcal{A}$, decide whether  $\mathcal{L(A)}$ is empty or not.
% For every NFA $\mathcal{A}$, there exist equivalent DFA $\mathcal{D}$, that is, $\mathcal{L(A)} = \mathcal{L(D)}$ (subset construction).
% Let's examine such DFA $\mathcal{D}$.
Since language of $\mathcal{A}$ is regular, then from pumping lemma we can "pump" words longer than some $N$ – pumping lemma constant.
That is, if $\mathcal{L(A)}$ has some word longer than $N$, then $\mathcal{L(A)}$ is infinite.
We just need to check every possible word $w$: $N < w \leq 2N$.
If any such word is accepted by $\mathcal{A}$, then $\mathcal{L(A)}$ is infinite.
We don't need to check words longer of $2N$: if a word is longer than $2N$, then from PL, we can iterative reduce its length, so $w$ has length shorter than $2N$.

How to choose $N$? We know that for every NFA $\mathcal{A}$, there exist equivalent DFA $\mathcal{D}$, that is, $\mathcal{L(A)} = \mathcal{L(D)}$ (subset construction).
We don't need to construct such DFA.
All we know is that, $\mathcal{D}$ might have exponentially more states than $\mathcal{A}$.
Take $N = |\Sigma|^{|Q_D|} + 42$, where $Q_D$ is set of $\mathcal{D}$'s states.